Regionally Additive Models: Explainable-by-design models minimizing feature interactions

### 2023-06-18 report/report.tex
continued the example
finished the example

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{geometry}[margin=1in]
\usepackage{subcaption}

@@

\usepackage{natbib}
\bibliographystyle{plainnat}
% set path for bibliography

@@

>>>>

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{geometry}[margin=1in]
\usepackage{subcaption}
\usepackage{array}

@@

\usepackage{natbib}
\usepackage{multicol}
\bibliographystyle{plainnat}
% set path for bibliography

@@

We evaluate the proposed approach on two typical tabular datasets, namely the Bike-Sharing Dataset~\citep{misc_bike_sharing_dataset_275}
and the California Housing Dataset~\citep{pace1997sparse}.

\paragraph{Bike-Sharing Dataset}

The Bike-Sharing dataset contains the hourly bike rentals in Washington DC. over the period 2011 and 2012.
The dataset contains 14 features and we select the 11 features that are relevant to the prediction task.
Most of the features are measurements of the environmental conditions, like,
month, hour, temperature, humidity, windspeed, while some others containt information about the
day-type, e.g. if it is working-day.
The target value \( Y_{\mathtt{count}}\) is the bike rentals per hour, which has mean value
\(\mu_{\mathtt{count}} = 189\) and standard deviation \(\sigma_{\mathtt{count}} = 181\).

% Training - Evaluation - Region Extraction
As black-box model, we train a deep fully-connected Neural Network with 6 hidden layers,
for \(20\) epochs using the Adam optimizer with learning rate 0.001, that
leads a root mean squared error of \( 0.38 \cdot 181 \approx 69\) counts on the test set.

We extract the subregions, setting the maximum depth of the tree to \(T=3\).
The only feature that is split is $X_{\mathtt{hour}}$, which is split into $X_{\mathtt{hour}} | \when{X_{\mathtt{workingday}} \neq 1}$ and $X_{\mathtt{hour}} \when{X_{\mathtt{workingday} = 1}}$.
As we observe in Figure (add citation), it is confirmed

\begin{figure}[htbp]
    \centering
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bike_rentals_gam}
        \caption{\(f(X_{\mathtt{hour}})\)}
        \label{subfig:bike_rentals_gam}
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bike_rentals_ram_1}
        \caption{\(f(X_{\mathtt{hour}}) \when{X_{\mathtt{workingday}} \neq 1}\)}
        \label{subfig:bike_rentals_regional_gam_1}
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bike_rentals_ram_2}
        \caption{\(f(X_{\mathtt{hour}}) \when{X_{\mathtt{workingday}} = 1}\)}
        \label{subfig:bike_rentals_regional_gam_2}
    \end{subfigure}
    \caption{Caption for the entire figure}
    \label{fig:bike_sharing}
\end{figure}

\paragraph{California Housing Dataset}
The California Housing is a largely-studied dataset with approximately
\(20000\) training instances, making it appropriate for robust
approximation with large \(K\).  The dataset contains \(D=8\)
numerical features with characteristics about the building blocks of
California, e.g., latitude, longitude, population of the block or
median age of houses in the block.  The target variable is the median
value of the houses inside the block in dollars that ranges between
\([15, 500] \cdot 10^3\), with a mean value of
\(\mu_Y \approx 201 \cdot 10^3 \) and a standard deviation of
\(\sigma_Y \approx 110 \cdot 10^3\). We exclude instances with missing
or outlier values and we normalize all features to zero-mean and unit
standard deviation. We split the dataset into \(N_{tr} = 15639\)
training and \(N_{test} = 3910\) test examples (80/20 split) and we
fit a Neural Network with 3 hidden layers of 256, 128 and 36 units
respectively.  After 15 epochs using the Adam optimizer with learning
rate \(\eta = 0.02\), the model achieves a MAE of \(37 \cdot 10^3\)
dollars.

\begin{figure}[htbp]
    \centering
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/california_gam}
        \caption{\(f(X_{\mathtt{longitude}})\)}
        \label{subfig:california_gam}
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/california_ram_1}
        \caption{\(f(X_{\mathtt{longitude}}) \when{X_{\mathtt{latitude}} \leq 34.89}\)}
        \label{subfig:california_ram_1}
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/california_ram_2}
        \caption{\(f(X_{\mathtt{longitude}}) \when{X_{\mathtt{latitude}} > 34.89}\)}
        \label{subfig:california_ram_2}
    \end{subfigure}
    \caption{Caption for the entire figure}
    \label{fig:california_housing}
\end{figure}

\begin{table}[htbp]
  \centering
  \caption{Result Comparison}
  \label{tab:sample}
  \begin{tabular}{l|c|cccc}
      \hline
      Type & \textbf{Black-box} & \multicolumn{4}{c}{\textbf{x-by-design}} \\
      \hline
      \hline
      Interactions & \textbf{all orders} & \multicolumn{2}{c}{1\textsuperscript{st} order} & \multicolumn{2}{c}{2\textsuperscript{nd} order} \\
      \hline
      \hline
      Model & \textbf{DNN} & \textbf{GAM} & \textbf{RAM} & \textbf{GAM}$^2$ & \textbf{RAM}$^2$ \\
      \hline
      Training & 0.135 & 0.374 & 0.274 & 0.181 & 0.176 \\
      Test     & 0.254 & 0.549 & 0.430 & 0.298 & 0.278 \\
      \hline
      Training & 0.282 & 0.322 & 0.314 & 0.278 & 0.265 \\
      Test     & 0.369 & 0.600 & 0.535 & 0.554 & 0.514 \\
  \end{tabular}
\end{table}
